---
title: 'Campbell sc-RNA seq - preprocess'
author: "Pers lab"
date: "`r Sys.time()`" 
output:
  html_notebook: 
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup

## load packages 

```{r}

library("here")
library("Seurat")
library("dplyr")
library("Matrix")
library("cowplot")
library("parallel")

```

## source functions

```{r}
source(here("perslab-sc-library", "utility_functions.R"))
source(here("perslab-sc-library", "functions_sc.R"))
```

## Set options

```{r}
options(stringsAsFactors = F, use="pairwise.complete.obs", warn=1)
```

## Set constants

```{r}
dir_project = "/projects/jonatan/tmp-holst-hsl/"
dir_plots = paste0(dir_project,"plots/")
dir_tables = paste0(dir_project,"tables/")
dir_log = paste0(dir_project,"log/")

prefix_data = "campbell"
flag_date = substr(gsub("-","",as.character(Sys.Date())),3,1000)
prefix_out <- "seurat_1"

```

# Download data

```{r}
download_url.data <- "ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE93nnn/GSE93374/suppl/GSE93374_Merged_all_020816_DGE.txt.gz"

download_url.metadata <- "ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE93nnn/GSE93374/suppl/GSE93374_cell_metadata.txt.gz"
path.data = here("data","GSE93374_Merged_all_020816_DGE.txt.gz")
path.metadata <- here("data", "GSE93374_cell_metadata.txt.gz")

```

```{r}
if (!file.exists(path.data)) {
 download.file(download_url.data, destfile=path.data) 
}

if (!file.exists(path.metadata)) {
 download.file(download_url.metadata, destfile=path.metadata) 
}
```

## Load sequencer data and do all QC

expression data

```{r}
# load data
seurat_object.data <- read.csv(gzfile(path.data),sep="\t")
```

metadata 

```{r} 

campMeta <- read.delim(path.metadata, quote = "", stringsAsFactors = F)
# intialize Seurat object
# keep all genes expressed in => 5 cells and all cells with at least 200 detected genes
rownames(campMeta) <- campMeta[["X1.ID"]]
campMeta <- campMeta[match(colnames(seurat_object.data), rownames(campMeta)),]

```

clear up non-alphanumeric characters in metadata

```{r}
campMeta[] <- apply(X=campMeta, MARGIN = 2, FUN = function(eachCol) {
  gsub("/|\\(|\\)","_", eachCol)
})

```

remove NA column 

```{r}
campMeta <- campMeta[,-12]
```

create Seurat object

```{r}
seurat_obj <- CreateSeuratObject(counts = seurat_object.data,
                             project = "ARCUATE_NUCLEUS",
                             min.cells=5,
                             meta.data = campMeta)

```

# add percent mito and ribo to metadata

```{r}
# Find mito + ribo gene percentages and nUMI and nGene and append to cells as meta data
mito.genes <- grep(pattern = "^mt-", x = rownames(x = seurat_obj), value = TRUE)
ribo.genes <- grep(pattern = "^Rp[sl][[:digit:]]", x = rownames(x = seurat_obj), value = TRUE)

percent.mito <- Matrix::colSums(seurat_obj@assays$RNA@counts[mito.genes, ])/Matrix::colSums(seurat_obj@assays$RNA@counts)
percent.ribo <- Matrix::colSums(seurat_obj@assays$RNA@counts[ribo.genes, ])/Matrix::colSums(seurat_obj@assays$RNA@counts)

seurat_object <- AddMetaData(object = seurat_obj, metadata = percent.mito, col.name = "percent.mito")
seurat_object <- AddMetaData(object = seurat_obj, metadata = percent.ribo, col.name = "percent.ribo")

```

Filter out cells whose gene reads are outside an appropriate range 

```{r}
seurat_obj <- subset(x = seurat_obj, 
                     subset = nFeature_RNA > 200 & nFeature_RNA < 10000 & percent.mito < 0.25)

```

Normalize the data

```{r}

seurat_object <- NormalizeData(object = seurat_object, normalization.method = "LogNormalize")

```

Find variable genes  

```{r}

seurat_obj <- FindVariableFeatures(object = seurat_obj,
                                   nfeatures = 2000,
                                   selection.method="vst", 
                                   verbose=T)

length(VariableFeatures(seurat_obj)) # Check how many genes are variable

```

Scale the data and regress out confounders 

```{r}

seurat_object <- ScaleData(object = seurat_object, 
                           vars.to.regress = c("nCount_RNA", "percent.mito"), 
                           features=VariableFeatures(seurat_object),
                           block.size=15000,
                           min.cells.to.block=5000,
                           verbose=T)
```

Run PCA

```{r}

seurat_object <- RunPCA(object = seurat_object, 
                        pc.genes = VariableFeatures(seurat_object), 
                        pcs.compute=50, 
                        do.print = F)#, pcs.print = 1:5, genes.print = 5)

# Elbow plot --> draw your cutoff where there is a clear elbow in the graph
PCElbowPlot(object = seurat_object)
pc_num = 9

seurat_object <- RunTSNE(object = seurat_object, dims.use = 1:pc_num, do.fast = TRUE)

```

## Reorganise annotations

```{r}
df_metadata <- data.frame(clust_all_full = as.character(seurat_obj$X7.clust_all), 
                          clust_sub_full = as.character(seurat_obj$X10.clust_neurons))

```

### subclusters

```{r}

for (cellID in unique(df_metadata$clust_all_full)) {
   if (length(grep(cellID, seurat_obj$All.Cell.Clusters, value=T))) {
    df_metadata$clust_all_full[df_metadata$clust_all_full==cellID] <- grep(cellID, seurat_obj$All.Cell.Clusters, value=T)
   }
} 

table(df_metadata$clust_all_full)
```

```{r}
for (cellID in unique(df_metadata$clust_sub_full)) {
  if (cellID=="miss") next
  if (length(grep(cellID, seurat_obj$Neuron.Subclusters, value=T))) {
    df_metadata$clust_sub_full[df_metadata$clust_sub_full==cellID] <- grep(cellID, seurat_obj$Neuron.Subclusters, value=T)
  }
}

table(df_metadata$clust_sub_full)
```

Add the full annotations to the n03, n07, n08 and n27 celltypes

```{r}
for (i in 1:4) {
  cellID <- c("n03", "n07", "n08", "n27")
  cellIDFull <- c("n03.Th_SSt", "n07.Arx_Nr5a2", "n08.Th_Slc6a3", "n27.Tbx19")
  df_metadata$clust_sub_full[df_metadata$clust_sub_full==cellID[i]] <- cellIDFull[i]
}

table(df_metadata$clust_sub_full)
```

```{r}
rownames(df_metadata) <- rownames(seurat_obj@meta.data)
seurat_obj<-AddMetaData(object=seurat_obj, metadata=df_metadata)
```

write out file to disk 

```{r}
saveRDS(seurat_obj, file = here(data, "campbell_all_clust_full.RDS.gz"), compress="gzip")
```

